\bf Tearing Mode Dynamics in Tokamak Plasmas

### 2021-12-04 Documentation/TearingModeTheory.tex
More corrections

According to the previous discussion, it is plausible that all terms appearing in Equation~(\ref{e176}) are of approximately equal
magnitude when the inertial and viscous terms are evaluated with the zeroth-order ion flow, which can be written
\begin{equation}
{\bf V}_i = r\,{\mit\Omega}_\theta(r,t)\,{\bf e}_\theta + R_0\,{\mit\Omega}_\varphi(r,t)\,{\bf e}_z.
\end{equation}
Here, ${\mit\Omega}_\theta(r,t)$ and ${\mit\Omega}_\varphi(r,t)$ are the ion poloidal and toroidal angular velocity
profiles, respectively. When $\{ r\,\nabla\cdot \bpi_i\cdot{\bf e}_\theta\}$ is evaluated with  the ion parallel viscosity tensor, and the zeroth-order
ion flow, it gives rise to a term that acts to relax the ion poloidal angular velocity profile toward a so-called {\em neoclassical
profile}\/ determined by ion density and temperature gradients (Stix 1973; Hirshman \& Sigmar 1981). In principle, the
contribution of the ion perpendicular viscosity tensor to $\{ r\,\nabla\cdot \bpi_i\cdot{\bf e}_\theta\}$ 
is ${\cal O}(\zeta_i^{\,2})$ smaller than the contribution of the ion parallel viscosity tensor. (See Section~\ref{sbrag}.) In practice, the magnitude of the ion perpendicular
viscosity tensor is greatly enhanced above the value predicted by the Braginskii equations by the action of small-scale
plasma turbulence (Wesson 2011). Hence, it is reasonable to include the contribution of the enhanced perpendicular
viscosity tensor to $\{ r\,\nabla\cdot \bpi_i\cdot{\bf e}_\theta\}$. Our model form for $\{ r\,\nabla\cdot \bpi_i\cdot{\bf e}_\theta\}$ is written
\begin{equation}\label{e178}
\{ r\,\nabla\cdot \bpi_i\cdot{\bf e}_\theta\} = 4\pi^2\,R_0\left[ \frac{\rho}{\tau_\theta}\,r^{\,3}\,({\mit\Omega}_\theta-{\mit\Omega}_{\theta\,{\rm nc}}) -\frac{\partial}{\partial r}\!\left(\mu_\perp\,r^{\,3}\,\frac{\partial{\mit\Omega}_\theta}{\partial r}\right)\right].
\end{equation}
Here, $\rho(r)$ is the equilibrium plasma mass density profile, $\tau_\theta(r)$ is the {\em poloidal flow-damping time}\/ [which
is generally of order $(a/q\,R_0)^2\,\tau_i$], ${\mit\Omega}_{\theta\,{\rm nc}}(r)$ is the neoclassical poloidal angular velocity profile, and
$\mu_\perp(r)$ is a phenomenological perpendicular viscosity due to plasma turbulence.

When evaluated with the zeroth-order ion flow, Equation~(\ref{e176}) reduces to 
\begin{equation}\label{e179}
4\pi^2\,R_0\left[\rho\,r^{\,3}\,\frac{\partial {\mit\Omega}_\theta}{\partial t}+ \frac{\rho}{\tau_\theta}\,r^{\,3}\,({\mit\Omega}_\theta-{\mit\Omega}_{\theta\,{\rm nc}}) -\frac{\partial}{\partial r}\!\left(\mu_\perp\,r^{\,3}\,\frac{\partial{\mit\Omega}_\theta}{\partial r}\right)\right]
=T_{\theta\,s}\,\delta(r-r_s)+ S_\theta,
\end{equation}
where use has been made of Equation~(\ref{e178}). 
Here, we have added a source term, $S_\theta(r)$, to the equation in order to account for the equilibrium poloidal flow. Let ${\mit\Omega}_{\theta\,0}(r)$ be the 
unperturbed (by the tearing mode) ion poloidal angular velocity profile. It follows that 
\begin{equation}
S_\theta = 4\pi^2\,R_0\left[ \frac{\rho}{\tau_\theta}\,r^{\,3}\,({\mit\Omega}_{\theta\,0}-{\mit\Omega}_{\theta\,{\rm nc}}) -\frac{d}{d r}\!\left(\mu_\perp\,r^{\,3}\,\frac{d{\mit\Omega}_{\theta\,0}}{d r}\right)\right].
\end{equation}
Hence, writing
\begin{equation}
{\mit\Omega}_\theta(r,t)= {\mit\Omega}_{\theta\,0}(r) + {\mit\Delta\Omega}_\theta(r,t),
\end{equation}
where ${\mit\Delta\Omega}_{\theta}(r,t)$ is the  modification to the ion poloidal angular velocity profile
induced by the poloidal electromagnetic torque that develops at the resonant surface, Equation~(\ref{e179})
yields 
\begin{equation}
4\pi^2\,R_0\left[\rho\,r^{\,3}\,\frac{\partial {\mit\Delta\Omega}_\theta}{\partial t}+ \frac{\rho}{\tau_\theta}\,r^{\,3}\,{\mit\Delta\Omega}_\theta-\frac{\partial}{\partial r}\!\left(\mu_\perp\,r^{\,3}\,\frac{\partial{\mit\Delta\Omega}_\theta}{\partial r}\right)\right]
=T_{\theta\,s}\,\delta(r-r_s).
\end{equation}
There is one further refinement that we can make to the previous equation. It turns out that neoclassical poloidal
flow damping gives rise to an enhancement of poloidal ion inertia by a factor of $1+2\,q^{\,2}$ (Hirshman 1978). Hence,
our final version of the perturbed plasma poloidal angular equation of motion becomes 
\begin{equation}\label{e182}
4\pi^2\,R_0\left[(1+2\,q^{\,2})\,\rho\,r^{\,3}\,\frac{\partial {\mit\Delta\Omega}_\theta}{\partial t}+ \frac{\rho}{\tau_\theta}\,r^{\,3}\,{\mit\Delta\Omega}_\theta-\frac{\partial}{\partial r}\!\left(\mu_\perp\,r^{\,3}\,\frac{\partial{\mit\Delta\Omega}_\theta}{\partial r}\right)\right]
=T_{\theta\,s}\,\delta(r-r_s).
\end{equation}

@@

\subsection{Solution of Plasma Angular Equations of Motion}
In most situations of interest, the perturbed angular velocity profiles, ${\mit\Delta\Omega}_\theta(r,t)$ and ${\mit\Delta\Omega}_\varphi(r,t)$, are
localized in the vicinity of the resonant surface (Fitzpatrick 1993). Hence, it is reasonable to express  the perturbed angular
equations of motion, (\ref{e182}) and (\ref{e185}), in the simplified  forms
\begin{align}\label{e188}
4\pi^2\,R_0\left[(1+2\,q_s^{\,2})\,\rho_s\,r^{\,3}\,\frac{\partial {\mit\Delta\Omega}_\theta}{\partial t}+ \frac{\rho_s}{\tau_{\theta\,s}}\,r^{\,3}\,{\mit\Delta\Omega}_\theta-\mu_{\perp\,s}\,\frac{\partial}{\partial r}\!\left(r^{\,3}\,\frac{\partial{\mit\Delta\Omega}_\theta}{\partial r}\right)\right]
&=T_{\theta\,s}\,\delta(r-r_s),\\[0.5ex]
4\pi^2\,R_0^{\,3}\left[\rho_s\,r\,\frac{\partial {\mit\Delta\Omega}_\varphi}{\partial t}-\mu_{\perp\,s}\,\frac{\partial}{\partial r}\!\left(r\,\frac{\partial{\mit\Delta\Omega}_\varphi}{\partial r}\right)\right]
=T_{\varphi\,s}\,\delta(r-r_s),\label{e189}
\end{align}
where $q_s=q(r_s)$, $\rho_s=\rho(r_s)$, $\tau_{\theta\,s}=\tau_\theta(r_s)$, and $\mu_{\perp\,s}=\mu_\perp(r_s)$.

@@

\subsection{Derivation of Linear Layer Equations}\label{s5.3}
Suppose that the system is perturbed by a tearing mode with $m$ periods in the poloidal direction, and
$n$ periods in the toroidal direction. In accordance with Equations~(\ref{e287})--(\ref{e290}), and (\ref{e292})--(\ref{e293}), we can write
\begin{align}\label{e295}
\hat{\psi}(\hat{x},\zeta,\hat{t}) &= \frac{\hat{x}^{\,2} }{2\,\hat{L}_s}+ \tilde{\psi}(\hat{x})\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},\\[0.5ex]
\phi(\hat{x},\zeta,\hat{t}) &=-\hat{V}_E\,\hat{x}+ \tilde{\phi}(\hat{x})\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},\\[0.5ex]
N(\hat{x},\zeta,\hat{t}) &= -\hat{V}_\ast\,\hat{x}+\left(\frac{1+\tau}{\tau}\right)\!\tilde{N}(\hat{x})\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},\\[0.5ex]
V(\hat{x},\zeta,\hat{t}) &= \hat{V}_\parallel + \left(\frac{1+\tau}{\tau}\right)\!\tilde{V}(\hat{x})\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},\\[0.5ex]
J(\hat{x},\zeta,\hat{t}) &=-\left(\frac{2}{s_s}-1\right)\!\frac{1}{\hat{L}_s}+ \hat{\nabla}^2\tilde{\psi}(\hat{x})\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},\\[0.5ex]
U(\hat{x},\zeta,\hat{t}) &=  \hat{\nabla}^2\tilde{\phi}(\hat{x})\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},\label{e300}
\end{align}
where $\hat{t}=V_A\,t/r_s$, $\hat{\omega}=r_s\,\omega/V_A$, and $\omega$ is the natural frequency of the tearing mode in the laboratory frame. 
Substituting Equations~(\ref{e294})--(\ref{e300}) into the four-field model, (\ref{e264})--(\ref{e269}), and only retaining terms that
are first order in perturbed quantities, we obtain the following set of linear layer equations:
\begin{align}\label{e301}
-{\rm i}\,[\omega-\omega_E -(1+\lambda_e)\,\omega_{\ast\,e}]\,\tau_H\,\tilde{\psi} &= -{\rm i} \,\hat{x}\,[\tilde{\phi}-(1+\lambda_e)\,\tilde{N}] + S^{-1}\,\hat{\nabla}^2\tilde{\psi},\\[0.5ex]
-{\rm i}\,(\omega-\omega_E-\omega_{\ast\,i})\,\tau_H\,\hat{\nabla}^2\tilde{\phi} &= -{\rm i}\,\hat{x}\,\hat{\nabla}^2\tilde{\psi}+S^{-1}\,P_\varphi\,\hat{\nabla}^4\!\left(\tilde{\phi} + \frac{\tilde{N}}{\tau}\right),\\[0.5ex]
-{\rm i}\,(\omega-\omega_E)\,\tau_H\,\tilde{N} &= -{\rm i}\,\omega_{\ast\,e}\,\tau_H\,\tilde{\phi}-{\rm i}\,c_\beta^{\,2}\,\hat{x}\,\tilde{V} - {\rm i}\!\left(\frac{\tau}{1+\tau}\right)\hat{d}_\beta^{\,2}\,\hat{x}\,\hat{\nabla}^2\tilde{\psi} \\[0.5ex]
&\phantom{=}
+S^{-1}\,P_\parallel\!\left(m\,\omega_{\ast\,e}\,\tau_H\,\hat{x}\,\tilde{\psi}-\frac{m}{\hat{L}_s}\,\hat{x}^2\,\tilde{N}\right)
+ S^{-1}\,P_\perp\,\hat{\nabla}_\perp^2 \tilde{N},\nonumber\\[0.5ex]
-{\rm i}\,(\omega-\omega_E)\,\tau_H\,\tilde{V} &= {\rm i}\,\omega_{\ast\,e}\,\tau_H\,\tilde{\psi} - {\rm i}\,\hat{x}\,\tilde{N}
+ S^{-1}\,P_\varphi\,\hat{\nabla}^2\tilde{V}.\label{e304}
\end{align}
Here, 
\begin{equation}\label{e313a}
\tau_H = \frac{L_s}{m\,V_A}
\end{equation}
is the {\em hydromagnetic time}, 
\begin{equation}
\omega_E = \left(\frac{m}{r_s}\right)V_E(r_s)
\end{equation}
  the 
{\em E-cross-B frequency}, 
\begin{equation}
\omega_{\ast\,e} = -\left(\frac{\tau}{1+\tau}\right)\left(\frac{m}{r_s}\right)V_\ast(r_s)
\end{equation}
  the {\em electron diamagnetic
frequency},
\begin{align}\label{omegai}
  \omega_{\ast\,i} &= \left(\frac{1}{1+\tau}\right)\left(\frac{m}{r_s}\right)V_\ast(r_s)
  \end{align}
    the {\em ion diamagnetic
frequency}, 
\begin{equation}
S=\frac{\tau_R}{\tau_H}
\end{equation}
 the {\em Lundquist number}, 
\begin{equation}\label{e317a}
\tau_R = \frac{\mu_0\,r_s^{\,2}}{\eta_\parallel}
\end{equation}
 the
{\em resistive diffusion time}, 
\begin{equation}\label{e318a}
\tau_\varphi
= \frac{r_s^{\,2}}{\chi_\varphi}=\frac{r_s^{\,2}\,n_0\,m_i}{\mu_\perp(r_s)}
\end{equation} 
the {\em toroidal momentum confinement time}, and 
\begin{equation}
\tau_\perp = \frac{r_s^{\,2}}{D_\perp}
\end{equation}
 the {\em particle/energy confinement time}. Furthermore, $P_\varphi = \tau_R/\tau_\varphi$, $P_\perp = \tau_R/\tau_\perp$, 
$P_\parallel=\tau_R/\tau_\parallel$, and $\tau_\parallel =r_s^{\,2}/D_\parallel$.

@@

Making use of Equations~(\ref{e193}), (\ref{e194}), (\ref{e205}) and (\ref{e206}) (with $d/dt=0$), (\ref{e383}), (\ref{e389}), (\ref{e392}),
and (\ref{e393}), we arrive at the following torque balance equation (Fitzpatrick 1993; Fitzpatrick 1998):
\begin{equation}\label{e394}
T_{\rm VS}(v)= T_{\rm EM}(v),
\end{equation}
where 
\begin{align}
v&=\frac{\omega_{s\,0}+{\mit\Delta}\omega_E}{\omega_{s\,0}},\\[0.5ex]
T_{\rm VS}(v) &= 1- v,\\[0.5ex]
T_{\rm EM}(v) &= \frac{1}{4}\,\frac{v}{\zeta_s^{\,2}+v^2}
\left(\frac{|\tilde{I}_c|}{\tilde{I}_{c\,{\rm crit}}}\right)^2,\\[0.5ex]
\zeta_s &= \frac{(-{\mit\Delta}_{nw})}{\tau_s\,\omega_{s\,0}}\ll 1,\\[0.5ex]
\tilde{I}_{c\,{\rm crit}}&= (\epsilon_s\,s_s)\,(\omega_{s\,0}\,\tau_H)\left/
\left[\frac{q_s}{\epsilon_s}\left(\frac{\sqrt{\tau_\varphi\,\tau_i/\mu_{00}^i}}{\tau_s}\right)+2\left(\frac{\tau_\varphi}{\tau_s}\right) \ln\!\left(\frac{a}{r_s}\right)\right]^{1/2}\right..\label{e420}
\end{align}
Here, use has been made of the identities
\begin{align}\label{e417a}
\lim_{\epsilon\rightarrow 0}\sum_{p=1,\infty} \frac{\sqrt{\epsilon}\,[J_1(j_{1\,p}\,x)]^{2}}{[J_2(j_{1\,p})]^2\,(1+\epsilon\,j_{1\,p}^2)}
&=\frac{1}{4\,x},\\[0.5ex]
\sum_{p=1,\infty} \frac{[J_0(j_{0\,p}\,x)]^2}{[J_1(j_{0\,p}\,x)]^2\,j_{0\,p}^{\,2}} =\frac{1}{2}\,\ln\!\left(\frac{1}{x}\right).\label{e418a}
\end{align}
We have also written 
\begin{equation}
\frac{1}{\tau_{\theta\,s}} = \left(\frac{q_s}{\epsilon_s}\right)^2\,\left(\frac{\mu_{00}^i}{\tau_{i}}\right),
\end{equation}
 where
$\tau_i$ is the ion collision time and $\mu_{00}^i$ is a dimensionless neoclassical viscosity coefficient defined in Fitzpatrick \& Nelson 2020. Both quantities are evaluated at the resonant surface. Furthermore,
 we have assumed that $\tau_i/(\tau_\varphi\,\mu_{00}^i)\ll (q_s/\epsilon_s)^2$ (this
turns out to be  
a very good approximation because the ion collision time, $\tau_i$, is much less than the momentum confinement
time, $\tau_\varphi$, in conventional tokamak plasmas).

@@

Finally, the criterion for the validity of the theory presented in this section is that the width of the magnetic island
chain driven at the resonant surface, just prior to mode penetration, is much less than the linear layer width (so that the resonant
plasma response is governed by linear physics). This is the case provided that (Fitzpatrick 1998)
\begin{equation}
\hat{\delta}\gg 4\,S^{-2/5}\left(\frac{q_s}{\epsilon_s}\right)^{2/5}\left[\frac{q_s}{\epsilon_s}\left(\frac{\sqrt{\tau_\varphi\,\tau_i/\mu_{00}^i}}{\tau_R}\right)+2\left(\frac{\tau_\varphi}{\tau_R}\right) \ln\!\left(\frac{a}{r_s}\right)\right]^{-1/5}.
\end{equation}

>>>>

According to the previous discussion, it is plausible that all terms appearing in Equation~(\ref{e176}) are of approximately equal
magnitude when the inertial and viscous terms are evaluated with the zeroth-order ion flow, which can be written
\begin{equation}
{\bf V}_i = r\,{\mit\Omega}_\theta(r,t)\,{\bf e}_\theta + R_0\,{\mit\Omega}_\varphi(r,t)\,{\bf e}_z.
\end{equation}
Here, ${\mit\Omega}_\theta(r,t)$ and ${\mit\Omega}_\varphi(r,t)$ are the ion poloidal and toroidal angular velocity
profiles, respectively. When $\{ r\,\nabla\cdot \bpi_i\cdot{\bf e}_\theta\}$ is evaluated with  the ion parallel viscosity tensor, and the zeroth-order
ion flow, it gives rise to a term that acts to relax the ion poloidal angular velocity profile toward a so-called {\em neoclassical
profile}\/ determined by ion density and temperature gradients (Stix 1973; Hirshman \& Sigmar 1981). In principle, the
contribution of the ion perpendicular viscosity tensor to $\{ r\,\nabla\cdot \bpi_i\cdot{\bf e}_\theta\}$ 
is ${\cal O}(\zeta_i^{\,2})$ smaller than the contribution of the ion parallel viscosity tensor. (See Section~\ref{sbrag}.) In practice, the magnitude of the ion perpendicular
viscosity tensor is greatly enhanced above the value predicted by the Braginskii equations by the action of small-scale
plasma turbulence (Wesson 2011). Hence, it is reasonable to include the contribution of the enhanced perpendicular
viscosity tensor to $\{ r\,\nabla\cdot \bpi_i\cdot{\bf e}_\theta\}$. Our model form for $\{ r\,\nabla\cdot \bpi_i\cdot{\bf e}_\theta\}$ is written
\begin{equation}\label{e178}
\{ r\,\nabla\cdot \bpi_i\cdot{\bf e}_\theta\} = 4\pi^2\,R_0\left[ \frac{\rho}{\tau_\theta}\,r^3\,({\mit\Omega}_\theta-{\mit\Omega}_{\theta\,{\rm nc}}) -\frac{\partial}{\partial r}\!\left(\mu_\perp\,r^3\,\frac{\partial{\mit\Omega}_\theta}{\partial r}\right)\right].
\end{equation}
Here, $\rho(r)$ is the equilibrium plasma mass density profile, $\tau_\theta(r)$ is the {\em poloidal flow-damping time}\/ [which
is generally of order $(a/q\,R_0)^2\,\tau_i$], ${\mit\Omega}_{\theta\,{\rm nc}}(r)$ is the neoclassical poloidal angular velocity profile, and
$\mu_\perp(r)$ is a phenomenological perpendicular viscosity due to plasma turbulence.

When evaluated with the zeroth-order ion flow, Equation~(\ref{e176}) reduces to 
\begin{equation}\label{e179}
4\pi^2\,R_0\left[\rho\,r^3\,\frac{\partial {\mit\Omega}_\theta}{\partial t}+ \frac{\rho}{\tau_\theta}\,r^3\,({\mit\Omega}_\theta-{\mit\Omega}_{\theta\,{\rm nc}}) -\frac{\partial}{\partial r}\!\left(\mu_\perp\,r^3\,\frac{\partial{\mit\Omega}_\theta}{\partial r}\right)\right]
=T_{\theta\,s}\,\delta(r-r_s)+ S_\theta,
\end{equation}
where use has been made of Equation~(\ref{e178}). 
Here, we have added a source term, $S_\theta(r)$, to the equation in order to account for the equilibrium poloidal flow. Let ${\mit\Omega}_{\theta\,0}(r)$ be the 
unperturbed (by the tearing mode) ion poloidal angular velocity profile. It follows that 
\begin{equation}
S_\theta = 4\pi^2\,R_0\left[ \frac{\rho}{\tau_\theta}\,r^3\,({\mit\Omega}_{\theta\,0}-{\mit\Omega}_{\theta\,{\rm nc}}) -\frac{d}{d r}\!\left(\mu_\perp\,r^3\,\frac{d{\mit\Omega}_{\theta\,0}}{d r}\right)\right].
\end{equation}
Hence, writing
\begin{equation}
{\mit\Omega}_\theta(r,t)= {\mit\Omega}_{\theta\,0}(r) + {\mit\Delta\Omega}_\theta(r,t),
\end{equation}
where ${\mit\Delta\Omega}_{\theta}(r,t)$ is the  modification to the ion poloidal angular velocity profile
induced by the poloidal electromagnetic torque that develops at the resonant surface, Equation~(\ref{e179})
yields 
\begin{equation}
4\pi^2\,R_0\left[\rho\,r^3\,\frac{\partial {\mit\Delta\Omega}_\theta}{\partial t}+ \frac{\rho}{\tau_\theta}\,r^3\,{\mit\Delta\Omega}_\theta-\frac{\partial}{\partial r}\!\left(\mu_\perp\,r^3\,\frac{\partial{\mit\Delta\Omega}_\theta}{\partial r}\right)\right]
=T_{\theta\,s}\,\delta(r-r_s).
\end{equation}
There is one further refinement that we can make to the previous equation. It turns out that neoclassical poloidal
flow damping gives rise to an enhancement of poloidal ion inertia by a factor of $1+2\,q^2$ (Hirshman 1978). Hence,
our final version of the perturbed plasma poloidal angular equation of motion becomes 
\begin{equation}\label{e182}
4\pi^2\,R_0\left[(1+2\,q^2)\,\rho\,r^3\,\frac{\partial {\mit\Delta\Omega}_\theta}{\partial t}+ \frac{\rho}{\tau_\theta}\,r^3\,{\mit\Delta\Omega}_\theta-\frac{\partial}{\partial r}\!\left(\mu_\perp\,r^3\,\frac{\partial{\mit\Delta\Omega}_\theta}{\partial r}\right)\right]
=T_{\theta\,s}\,\delta(r-r_s).
\end{equation}

@@

\subsection{Solution of Plasma Angular Equations of Motion}
In most situations of interest, the perturbed angular velocity profiles, ${\mit\Delta\Omega}_\theta(r,t)$ and ${\mit\Delta\Omega}_\varphi(r,t)$, are
localized in the vicinity of the resonant surface (Fitzpatrick 1993). Hence, it is reasonable to express  the perturbed angular
equations of motion, (\ref{e182}) and (\ref{e185}), in the simplified  forms
\begin{align}\label{e188}
4\pi^2\,R_0\left[(1+2\,q_s^{\,2})\,\rho_s\,r^3\,\frac{\partial {\mit\Delta\Omega}_\theta}{\partial t}+ \frac{\rho_s}{\tau_{\theta\,s}}\,r^3\,{\mit\Delta\Omega}_\theta-\mu_{\perp\,s}\,\frac{\partial}{\partial r}\!\left(r^3\,\frac{\partial{\mit\Delta\Omega}_\theta}{\partial r}\right)\right]
&=T_{\theta\,s}\,\delta(r-r_s),\\[0.5ex]
4\pi^2\,R_0^{\,3}\left[\rho_s\,r\,\frac{\partial {\mit\Delta\Omega}_\varphi}{\partial t}-\mu_{\perp\,s}\,\frac{\partial}{\partial r}\!\left(r\,\frac{\partial{\mit\Delta\Omega}_\varphi}{\partial r}\right)\right]
=T_{\varphi\,s}\,\delta(r-r_s),\label{e189}
\end{align}
where $q_s=q(r_s)$, $\rho_s=\rho(r_s)$, $\tau_{\theta\,s}=\tau_\theta(r_s)$, and $\mu_{\perp\,s}=\mu_\perp(r_s)$.

@@

\subsection{Derivation of Linear Layer Equations}\label{s5.3}
Suppose that the system is perturbed by a tearing mode with $m$ periods in the poloidal direction, and
$n$ periods in the toroidal direction. In accordance with Equations~(\ref{e287})--(\ref{e290}), and (\ref{e292})--(\ref{e293}), we can write
\begin{align}\label{e295}
\hat{\psi}(\hat{x},\zeta,\hat{t}) &= \frac{\hat{x}^{\,2} }{2\,\hat{L}_s}+ \tilde{\psi}(\hat{x})\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},\\[0.5ex]
\phi(\hat{x},\zeta,\hat{t}) &=-\hat{V}_E\,\hat{x}+ \tilde{\phi}(\hat{x})\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},\\[0.5ex]
N(\hat{x},\zeta,\hat{t}) &= -\hat{V}_\ast\,\hat{x}+\left(\frac{1+\tau}{\tau}\right)\!\tilde{N}(\hat{x})\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},\\[0.5ex]
V(\hat{x},\zeta,\hat{t}) &= \hat{V}_\parallel + \left(\frac{1+\tau}{\tau}\right)\!\tilde{V}(\hat{x})\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},\\[0.5ex]
J(\hat{x},\zeta,\hat{t}) &=-\left(\frac{2}{s_s}-1\right)\!\frac{1}{\hat{L}_s}+ \hat{\nabla}^2\tilde{\psi}(\hat{x})\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},\\[0.5ex]
U(\hat{x},\zeta,\hat{t}) &=  \hat{\nabla}^2\tilde{\phi}(\hat{x})\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},\label{e300}
\end{align}
where $\hat{t}=V_A\,t/r_s$, $\hat{\omega}=r_s\,\omega/V_A$, and $\omega$ is the frequency of the tearing mode in the laboratory frame. 
Substituting Equations~(\ref{e294})--(\ref{e300}) into the four-field model, (\ref{e264})--(\ref{e269}), and only retaining terms that
are first order in perturbed quantities, we obtain the following set of linear layer equations:
\begin{align}\label{e301}
-{\rm i}\,[\omega-\omega_E -(1+\lambda_e)\,\omega_{\ast\,e}]\,\tau_H\,\tilde{\psi} &= -{\rm i} \,\hat{x}\,[\tilde{\phi}-(1+\lambda_e)\,\tilde{N}] + S^{-1}\,\hat{\nabla}^2\tilde{\psi},\\[0.5ex]
-{\rm i}\,(\omega-\omega_E-\omega_{\ast\,i})\,\tau_H\,\hat{\nabla}^2\tilde{\phi} &= -{\rm i}\,\hat{x}\,\hat{\nabla}^2\tilde{\psi}+S^{-1}\,P_\varphi\,\hat{\nabla}^4\!\left(\tilde{\phi} + \frac{\tilde{N}}{\tau}\right),\\[0.5ex]
-{\rm i}\,(\omega-\omega_E)\,\tau_H\,\tilde{N} &= -{\rm i}\,\omega_{\ast\,e}\,\tau_H\,\tilde{\phi}-{\rm i}\,c_\beta^{\,2}\,\hat{x}\,\tilde{V} - {\rm i}\!\left(\frac{\tau}{1+\tau}\right)\hat{d}_\beta^{\,2}\,\hat{x}\,\hat{\nabla}^2\tilde{\psi} \\[0.5ex]
&\phantom{=}
+S^{-1}\,P_\parallel\!\left(m\,\omega_{\ast\,e}\,\tau_H\,\hat{x}\,\tilde{\psi}-\frac{m}{\hat{L}_s}\,\hat{x}^2\,\tilde{N}\right)
+ S^{-1}\,P_\perp\,\hat{\nabla}_\perp^2 \tilde{N},\nonumber\\[0.5ex]
-{\rm i}\,(\omega-\omega_E)\,\tau_H\,\tilde{V} &= {\rm i}\,\omega_{\ast\,e}\,\tau_H\,\tilde{\psi} - {\rm i}\,\hat{x}\,\tilde{N}
+ S^{-1}\,P_\varphi\,\hat{\nabla}^2\tilde{V}.\label{e304}
\end{align}
Here, 
\begin{equation}\label{e313a}
\tau_H = \frac{L_s}{m\,V_A}
\end{equation}
is the {\em hydromagnetic time}, 
\begin{equation}
\omega_E = \left(\frac{m}{r_s}\right)V_E(r_s)
\end{equation}
  the 
{\em E-cross-B frequency}, 
\begin{equation}
\omega_{\ast\,e} = -\left(\frac{\tau}{1+\tau}\right)\left(\frac{m}{r_s}\right)V_\ast(r_s)
\end{equation}
  the {\em electron diamagnetic
frequency},
\begin{align}\label{omegai}
  \omega_{\ast\,i} &= \left(\frac{1}{1+\tau}\right)\left(\frac{m}{r_s}\right)V_\ast(r_s)
  \end{align}
    the {\em ion diamagnetic
frequency}, 
\begin{equation}
S=\frac{\tau_R}{\tau_H}
\end{equation}
 the {\em Lundquist number}, 
\begin{equation}\label{e317a}
\tau_R = \frac{\mu_0\,r_s^{\,2}}{\eta_\parallel}
\end{equation}
 the
{\em resistive diffusion time}, 
\begin{equation}\label{e318a}
\tau_\varphi
= \frac{r_s^{\,2}}{\chi_\varphi}=\frac{r_s^{\,2}\,n_0\,m_i}{\mu_\perp(r_s)}
\end{equation} 
the {\em toroidal momentum confinement time}, and 
\begin{equation}
\tau_\perp = \frac{r_s^{\,2}}{D_\perp}
\end{equation}
 the {\em particle/energy confinement time}. Furthermore, $P_\varphi = \tau_R/\tau_\varphi$, $P_\perp = \tau_R/\tau_\perp$, 
$P_\parallel=\tau_R/\tau_\parallel$, and $\tau_\parallel =r_s^{\,2}/D_\parallel$.

@@

Making use of Equations~(\ref{e193}), (\ref{e194}), (\ref{e205}) and (\ref{e206}) (with $d/dt=0$), (\ref{e383}), (\ref{e389}), (\ref{e392}),
and (\ref{e393}), we arrive at the following torque balance equation (Fitzpatrick 1993; Fitzpatrick 1998):
\begin{equation}\label{e394}
T_{\rm VS}(v)= T_{\rm EM}(v),
\end{equation}
where 
\begin{align}
v&=\frac{\omega_{s\,0}+{\mit\Delta}\omega_E}{\omega_{s\,0}},\\[0.5ex]
T_{\rm VS}(v) &= 1- v,\\[0.5ex]
T_{\rm EM}(v) &= \frac{1}{4}\,\frac{v}{\zeta_s^{2}+v^2}
\left(\frac{|\tilde{I}_c|}{\tilde{I}_{c\,{\rm crit}}}\right)^2,\\[0.5ex]
\zeta_s &= \frac{(-{\mit\Delta}_{nw})}{\tau_s\,\omega_{s\,0}}\ll 1,\\[0.5ex]
\tilde{I}_{c\,{\rm crit}}&=\epsilon_s\,s_s\left(\frac{\tau_s\,\tau_H\,\omega_{s\,0}}{\tau_V}\right),\label{e420}
\\[0.5ex]
\tau_V&= \frac{q_s}{\epsilon_s}\left(\frac{\tau_\varphi\,\tau_i}{\mu_{00}^i}\right)^{1/2}+ 2\ \ln\!\left(\frac{a}{r_s}\right)\tau_\varphi.\label{e421a}
\end{align}
Here, use has been made of the identities
\begin{align}\label{e417a}
\lim_{\epsilon\rightarrow 0}\sum_{p=1,\infty} \frac{\sqrt{\epsilon}\,[J_1(j_{1\,p}\,x)]^{2}}{[J_2(j_{1\,p})]^2\,(1+\epsilon\,j_{1\,p}^2)}
&=\frac{1}{4\,x},\\[0.5ex]
\sum_{p=1,\infty} \frac{[J_0(j_{0\,p}\,x)]^2}{[J_1(j_{0\,p}\,x)]^2\,j_{0\,p}^{\,2}} =\frac{1}{2}\,\ln\!\left(\frac{1}{x}\right).\label{e418a}
\end{align}
We have also written 
\begin{equation}
\frac{1}{\tau_{\theta\,s}} = \left(\frac{q_s}{\epsilon_s}\right)^2\,\left(\frac{\mu_{00}^i}{\tau_{i}}\right),
\end{equation}
 where
$\tau_i$ is the ion collision time and $\mu_{00}^i$ is a dimensionless neoclassical viscosity coefficient defined in Fitzpatrick \& Nelson 2020. Both quantities are evaluated at the resonant surface. Furthermore,
 we have assumed that $\tau_i/(\tau_\varphi\,\mu_{00}^i)\ll (q_s/\epsilon_s)^2$ (this
turns out to be  
a very good approximation because the ion collision time, $\tau_i$, is much less than the momentum confinement
time, $\tau_\varphi$, in conventional tokamak plasmas).

@@

Finally, the criterion for the validity of the theory presented in this section is that the width of the magnetic island
chain driven at the resonant surface, just prior to mode penetration, is much less than the linear layer width (so that the resonant
plasma response is governed by linear physics). This is the case provided that (Fitzpatrick 1998)
\begin{equation}
\hat{\delta}\gg 4\left(\frac{q_s}{\epsilon_s}\right)^{2/5}\left(\frac{\tau_H^{\,2}}{\tau_R\,\tau_V}\right)^{1/5}.
\end{equation}
